---
title: "BBets"
author: "Mikaela Elias"
output: html_document
date: "2024-09-26"
---


This assignment has three questions:  

1. in a class of 35 students, Dr. J bets $1 against student's $1  that at least 2 students have the same birthday, month, day.
Find the probability that at least two student have the same birthday.  If 26 students in the class take the bet how much is Dr. J expected to win/lose.   How much should Dr. J wager against the student's $1 to make this a fair bet?      

The probability of 2 students having the same birthday is 0.8152. Dr.J is expected to win 16.3904. To make this a fair bet Dr.J would have to put in $4.41.

2. Assume class size is 26 instead of 35.  With the same bet ($1 ot $1) what is the probability Dr. J wins the bet now.  If 20 of the 26 students make the bet, what is Dr. J's expected profit/loss?

The probability that Dr.J wins the bet now is 0.6066. Dr.J expected profit is $4.264. 

3. What size class (how many students) makes the $1 againt $1 a fair bet?

To make the bet fair there would need to be 23 students. 


```{r setup, include=FALSE}
# Load necessary libraries
library(ggplot2)
```

```{r}
# Function to simulate the birthday problem
simulate_birthday <- function(class_size = 35, trials = 100) {
  results <- numeric(trials)
  
  for (i in 1:trials) {
    birthdays <- sample(1:365, class_size, replace = TRUE) # Sample birthdays
    has_duplicate <- any(duplicated(birthdays)) # Check for duplicates
    results[i] <- ifelse(has_duplicate, 1, 0) # Store 1 if duplicate found, 0 otherwise
  }
  
  # Calculate the probability of at least one duplicate
  probability <- mean(results)
  return(probability)
}

# Run the simulation 100 times
set.seed(123) # Set seed for reproducibility
probabilities <- replicate(100, simulate_birthday())

# Print mean of probabilities
mean_prob <- mean(probabilities)
cat("Mean =", mean_prob, "\n")


```

```{r}
# Plot histogram of probabilities
ggplot(data.frame(probabilities), aes(x = probabilities)) +
  geom_histogram(bins = 10, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Probabilities",
       x = "Probability",
       y = "Frequency") +
  theme_minimal()


```
```{r}
# Calculate 95% confidence interval
ci <- quantile(probabilities, c(0.025, 0.975))
cat("95% Confidence Interval: [", ci[1], ", ", ci[2], "]\n")


```
```{r}
EV= mean_prob*26-(1-mean_prob)*26
print(paste("ev=",EV))
```

```{r}
print(paste('mean=',mean(probabilities)))
```
```{r}
FAIR=(mean_prob)/(1-mean_prob)
print(paste("fairbet=",FAIR))
```
#Question 2

```{r}
# Function to simulate the birthday problem
simulate_birthday <- function(class_size = 26, trials = 100) {
  results <- numeric(trials)
  
  for (i in 1:trials) {
    birthdays <- sample(1:365, class_size, replace = TRUE) # Sample birthdays
    has_duplicate <- any(duplicated(birthdays)) # Check for duplicates
    results[i] <- ifelse(has_duplicate, 1, 0) # Store 1 if duplicate found, 0 otherwise
  }
  
  # Calculate the probability of at least one duplicate
  probability <- mean(results)
  return(probability)
}

# Run the simulation 100 times
set.seed(123) # Set seed for reproducibility
probabilities <- replicate(100, simulate_birthday())

# Print mean of probabilities
mean_prob <- mean(probabilities)
cat("Mean =", mean_prob, "\n")


```
```{r}
EV= mean_prob*20-(1-mean_prob)*20
print(paste("ev=",EV))
```
#Question 3

```{r}
# Function to simulate the birthday problem
simulate_birthday <- function(class_size = 23, trials = 100) {
  results <- numeric(trials)
  
  for (i in 1:trials) {
    birthdays <- sample(1:365, class_size, replace = TRUE) # Sample birthdays
    has_duplicate <- any(duplicated(birthdays)) # Check for duplicates
    results[i] <- ifelse(has_duplicate, 1, 0) # Store 1 if duplicate found, 0 otherwise
  }
  
  # Calculate the probability of at least one duplicate
  probability <- mean(results)
  return(probability)
}

# Run the simulation 100 times
set.seed(123) # Set seed for reproducibility
probabilities <- replicate(100, simulate_birthday())

# Print mean of probabilities
mean_prob <- mean(probabilities)
cat("Mean =", mean_prob, "\n")


```
```{r}
EV= mean_prob*20-(1-mean_prob)*20
print(paste("ev=",EV))
```

